---
title: "JSCCA Vignette"
author: "Project"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
---

## Overview

This vignette demonstrates how to run the JSCCA implementation in
`jscca.R` using synthetic data.

The implementation follows the update/deflation scheme from:

- Dutta D, Sen A, Satagopan JM (2024), *Genetic Epidemiology*
- Paper: <https://pmc.ncbi.nlm.nih.gov/articles/PMC11589067/>

## 1) Load Functions

```{r setup}
source("../jscca.R")
set.seed(2026)
```

If your working directory is the project root, `../jscca.R` is correct when
knitting this file from `vignettes/`.

## 2) Simulate Data

```{r simulate}
n <- 150
p <- 60   # CNA features
q <- 80   # Methylation features
r <- 70   # Gene expression features

C <- matrix(rnorm(n * p), nrow = n, ncol = p)
M <- matrix(rnorm(n * q), nrow = n, ncol = q)
E <- matrix(rnorm(n * r), nrow = n, ncol = r)

# Add a weak shared latent structure to make signal visible
z <- scale(rnorm(n))
C[, 1:3] <- C[, 1:3] + 0.6 * z
M[, 5:8] <- M[, 5:8] + 0.6 * z
E[, 2:4] <- E[, 2:4] + 0.6 * z
```

## 3) Fit JSCCA (Hotelling Deflation)

```{r fit-hotelling}
fit_hotelling <- jscca_fit(
  C = C,
  M = M,
  E = E,
  ncomp = 3,
  lambda_u = 6,
  lambda_w = 8,
  lambda_v = 7,
  deflation = "hotelling",
  update_mode = "paper",
  standardize = TRUE,
  tol = 1e-6,
  max_iter = 500,
  verbose = TRUE
)

fit_hotelling$q_cm
fit_hotelling$q_me
fit_hotelling$converged
```

## 4) Fit JSCCA (Sub-Matrix Deflation by non-zero u and v)

This mode removes selected CNA and gene-expression variables between
components (while keeping methylation space intact).

```{r fit-submatrix}
fit_submatrix <- jscca_fit(
  C = C,
  M = M,
  E = E,
  ncomp = 3,
  lambda_u = 6,
  lambda_w = 8,
  lambda_v = 7,
  deflation = "submatrix_uv",
  update_mode = "paper",
  standardize = TRUE,
  tol = 1e-6,
  max_iter = 500,
  verbose = TRUE
)

fit_submatrix$q_cm
fit_submatrix$q_me
fit_submatrix$converged
```

## 5) Inspect Selected Variables

Selected indices are returned in original feature indexing.

```{r inspect}
fit_submatrix$selected_u
fit_submatrix$selected_w
fit_submatrix$selected_v
```

You can also inspect full loading vectors:

```{r loadings}
dim(fit_submatrix$U)  # p x ncomp
dim(fit_submatrix$W)  # q x ncomp
dim(fit_submatrix$V)  # r x ncomp
```

## 6) Notes on Tuning

- `lambda_u`, `lambda_w`, `lambda_v` control sparsity.
- Smaller lambda generally increases sparsity after thresholding.
- `ncomp` is the maximum extracted components; extraction may stop early if
  active submatrices become empty in `submatrix_uv` mode.
- `update_mode = "paper"` uses the appendix-style previous-iterate updates.
  `"alternating"` uses newest-available updates in each cycle.
